### 2016-07-17 version0.0/paper2.tex
finishing intro0.0
fig.s and table corrections

\chapter[SOM: classifying high $Z$ galaxies]{Clustering galaxy spectra at $0.5<z<1$: an unsupervised approach}
\defcitealias{Hossein12}{T12}
\defcitealias{Kinney96}{K96}
\defcitealias{Noll09}{N09}

@@

 \subsection{Kinney spectral model}
     \begin{figure}
        \centering
        \includegraphics[width=\textwidth]{../image_paper2/k96.jpg}
        \caption{\citet{Kinney96} spectra template for 12 types of galaxies from \citet{Hossein12} paper (Fig. 1 in \citet{Hossein12}). The type of each template is shown in each frame. Plots B, E, S0, Sa, Sb and SC show spectra that belong to the early types galaxies. Starburst galaxies spectra are indicated with SB 1 to 6. Higher numbers represent more intrinsic colour extinction.}
        \label{fig: k96}
    \end{figure}

@@

 \subsection{SED and properties of the sample galaxies} 
    \citetalias{Hossein12} selected 142 galaxies from the spectroscopic campaign of the ESO GOODS-South field~\citep{Vanzella05, Vanzella06, Vanzella08}.
    The 142 galaxies were selected based on the availability of photometry from HST/ACS, VLA/ISAAC, and {\it Spitzer}/MIPS and IRAC (10--13 filters with $\sim 0.4<\lambda<24~\mu$m in the observed frame).
   Data from these instruments was necessary in order to have a complete picture of stellar population and star formation rate. 
    For each galaxy, a robust spectroscopic redshift and photometric measurements from the GOODS-MUSIC catalogue \citep{Santini09} was available.
   \citetalias{Hossein12} matched the point spread function (PSF) of the photometric data with \citetalias{Kinney96} data and used the photometry as inputs to the Code Investigating GALaxy Emission ({\em CIGALE});~\citep[][hereafter N09]{Noll09} to generate the best-fit SED for each galaxy.
   % as well as some of the physical properties of the galaxies.
% PB 20160603: commented out following 3 lines as I don't think this fit well here. Could maybe go somewhere else.    
%    {\em CIGALE} is a valuable tool to investigate the properties of galaxies using ultraviolet to infrared wavebands.
%    It uses stellar populations, synthetic attenuation and dust emission models, spectral line templates, and active galactic nuclei optical spectral templates to model the SEDs of galaxies.
%    This code was successfully tested on data from 39 galaxies selected from the Spitzer Infrared Nearby Galaxy Survey (SINGS;~\citep{Kennicutt03}) by \citetalias{Noll09}.
    \citetalias{Hossein12} produced the best SED match, with wavelength interval of 910~\AA~to $\sim 80$~cm, for each galaxy\footnote{This wavelength interval is the default output of the {\em CIGALE} code}.
    Assuming decreasing SFR and visual attenuation ($\tau$) model, Salpeter initial mass function~\citep{Salpeter55}, and old stellar population with age of $\sim 10$~Gyr, they derived physical properties of the galaxies such as age and stellar mass.
    Some of these properties are shown in Tab.~\ref{tab: props}.
    In Sec.~\ref{sec: 1D}, we study these properties for each category.
    More details on creating SEDs and extracting information about galaxy properties using {\em CIGALE} can be found in \citetalias{Noll09} and \citetalias{Hossein12}.

\begin{table}
\caption{Description of the properties of \citet{Hossein12} galaxies; the output result of {\em CIGALE}}     
\label{tab: props}
\centering
\begin{tabular}{l l l}
\hline\hline
\noalign{\smallskip}
Par. & Unit & Description\\
\noalign{\smallskip}
\hline
\noalign{\smallskip}
$t_{\,\mathrm{oSP}}$ & Gyr & age of old SP model \\
$t_{\,\mathrm{ySP}}$ & Gyr & age of young SP model \\
$f_\mathrm{burst}$ & --- & mass fraction of \\
& & young single population (SP) model \\
\noalign{\smallskip}
$t_{\,\mathrm{D4000}}$ & Gyr & D4000-related age \\
\noalign{\smallskip}
$M_\mathrm{star}$ & M$_\odot$ & total stellar mass  \\
SFR & M$_\odot$/yr & instantaneous SFR  \\
$A_\mathrm{FUV}$ & mag & attenuation at 1500\,\AA{} \\
\noalign{\smallskip}
\hline
\end{tabular}
\end{table}

@@

\subsection{Creating self-organizing maps}
\label{sec: create_som}
     In order to create SOMs, we use {\sc matlab} neural network toolbox~\citep[NNT,][]{matlabtolbox}.
     %%Sr160611: Last time I got the following note from the MNRAS editor:
     %"Please note that computer software/programming lan- guages must be styled in SMALL CAPITAL LETTERS, according to journal style. Please check and correct this paper accordingly."
     % PB20160603: right, but I think the right way to do this is {\sc newsom}, not {\tiny NEWSOM} -- the distinction being that {\sc.. } will scale up or down with the global font size but {\tiny} will not.
     SOM in {\sc nnt} can be created by {\sc newsom} or {\sc selforgmap} library, both of which work in two phases, an ``ordering phase" and a ``tuning phase". 
     The first phase is called the ``ordering phase" and
     starts with maximum neighbourhood distance and an initial high learning factor (usually 0.9) is provided by the user. 
     The ordering phase continues for a requested number of iterations.
     During the iterations, the learning factor decreases to the tuning phase learning factor and the neighbourhood distance reaches that of the tuning phase as well.
     Both the learning factor and the neighbourhood of the tuning phase are set by the user. 
     The amount by which these two factors change in each iteration depends on the number of iterations.

     In the second, or ``tuning'' phase,
     the neighbourhood distance is kept at the user-defined minimum.
     The learning factor, however, decreases gradually.
     The gradual change in the leaning factor helps to fine-tune the topology results, leading to a more stable SOM. 
     To allow the fine tuning, the number of iterations in this phase must be much larger than the that of the ordering phase. %cite kohenen book?!
     We chose the number of epochs in the tuning phase to be 3 times the number of epochs in the ordering phase.

     To present our results, we use {\sc nnt}'s built-in plotting tool.
%     This tool is designed to show the distance between each cluster more clearly.
     Specifically, we use two of the plots in this tool: a hits map, which shows the number of times each neuron has became the winner (hits), and a distance map, which shows the distance between those neurons.
     In the maps, the purple hexagonal shapes represent the neurons. 
     The distances in a distance map are shown by the grey cycle colours:
     the darker the colour, the larger the distance between neurons.
     In the hit maps, neurons with zero hits are left empty.

@@

         \begin{figure}
            \begin{subfigure}[b]{0.5\textwidth}
                \centering
                \includegraphics[width=\textwidth]{../image_paper2/sample/sample2_dist.png}
            \end{subfigure}
            \hfill
            \begin{subfigure}[b]{0.5\textwidth}
                \includegraphics[width=\textwidth]{../image_paper2/sample/sample2_hits.png}
            \end{subfigure}
            \caption{SOM of the mock sample. In both plots, the axes show the position of the neurons. Hexagonal shapes represent the neurons. The top plot is a distance map. The grey cycle colours show the differences between the weights of each neuron with white being the minimum difference and black being the maximum one. The lower plot is a ``hits plot". It represents the number of samples in each neuron, where empty means zero hits. In this sample, the 27 galaxies are clustered in 4 groups, containing 8, 7, 6 and 6 galaxies respectively.}
            \label{fig: sample}
        \end{figure}

 To illustrate how self-organizing maps work, we create a mock sample of 27 galaxies.
% We create a mock sample of 27 galaxies to show how this method works, illustrate the results of SOMs, and how they can be analysed.(<- I don't like this sentence)
 The mock sample contains two attributes for each galaxy: the type (type 1 or type 2), and the star formation rate (high or low). 
 We generated a SOM of size $10 \times 10$ using the same initial values mentioned in Sec.~\ref{sec: create_som}.

@@

Using this method, as expected, we are able to divide the mock sample galaxies into 4 distinct groups: Type 1 with high SFR, type 1 with low SFR, type 2 with high SFR and type 2 with low SFR. 
The upper panel in the Fig.~\ref{fig: sample} clearly shows this division.
In that plot, the upper half belongs to high star forming galaxies, while the lower half belongs to low star forming galaxies.
The left half of the plot is where type 1 galaxies belong, and type 2 galaxies reside in the right side.
Different shades of gray show the border between regions.
The lower panel in Fig.~\ref{fig: sample} shows that only 4 neurons (out of 100) are occupied. 
8 galaxies are type 1 galaxies with high SFR, 7 are type 2 galaxies with high SFR, 6 are type 1 galaxies with low SFR and the other 6 are type 2 galaxies with low SFR. 
%Since the input data are chosen to not to be complex, each entry only could get either 0 or 1 as a type of galaxy and 0 or 0.5 as an indicator of  high or low SFR, only four of the neurons were filled.
We can conclude that although each galaxy had  the chance to occupy any of the neurons in the network, because of the similarity in the values of their attributes, they remained only in four groups.
This network is considered a training network and can be used to cluster any new dataset with similar entries.

@@

    In this section we show the results of the neural networks trained using the \citetalias{Kinney96} template spectra.
    Training with \citetalias{Kinney96} templates results in networks that have regions corresponding to galaxies of known morphological type. 
    The trained networks can then be used to categorize other galaxies.
    %In following we show how we trained data and used galaxies from \citetalias{Hossein12} to show how the clustering using trained networks works.

@@

    We started our analysis by creating 1D SOMs. 
    First, we created SOMs with only two neurons ($1\times2$ map), and then increased the number of neurons one at a time in the 1D case (Sec.~\ref{sec: 1D}).
    We generate 2D networks (Sec.~\ref{sec: 2D}),  again starting with the smallest possible number of neurons (4 neurons in a $2\times2$ map), increasing to 144 in a $12\times12$ map.    
    For each generated network, we compare the results with the \citetalias{Kinney96} categorization.
    We also use these networks to classify the \citetalias{Hossein12} galaxy sample, and compare this classification with that from the supervised networks in \citetalias{Hossein12}.

@@

            Figs.~\ref{fig: 1by2T} --\ref{fig: 1by22T} show the results of the training networks.
            As in to Fig.~\ref{fig: sample}, the upper part of the figures shows the neurons and their relative distances, $D_j$.
            As mentioned in Sec.~\ref{sec: method}, an increase in the darkness of colours between neurons represents an increase in relative distance between the neurons.
            The lower panels of the figures show the number of \citetalias{Kinney96} templates that are placed in each neuron. 
            \begin{figure}
                \begin{subfigure}[b]{0.5\textwidth}
                    \centering
                    \includegraphics[width=\textwidth]{../image_paper2/1d/dist_1_by_2.png}
                \end{subfigure}
                \hfill
                \begin{subfigure}[b]{0.5\textwidth}
                     \includegraphics[width=\textwidth]{../image_paper2/1d/hit_t_1_by_2.png}
                \end{subfigure}
                \caption{Results of training network in $1\times2$~grid. As in Fig.~\ref{fig: sample}, the upper panel is a distance map and the lower panel is a hit map. In this network, 5 of the templates from \citet{Kinney96} are categorized as quiescent galaxies and the rest are starbursts. Because of their strong emission lines, Sc galaxies are moved towards the starburst ones.}
                 \label{fig: 1by2T}
            \end{figure}

@@

            Fig.~\ref{fig: 1by3T} shows the results of the training in a 1$\times$3 network.
            In these plots we force the galaxies to be categorize in a maximum 3 groups. 
            %If the galaxies in Fig.~\ref{fig: 1by2T} were very similar to each other or were exactly the same, they still wanted to be divided into two groups and left the middle node empty(this sentence does not make much sense. I kind of think I know what you're trying to say (because once I translate this sentence to farsi word by word, it kind of comes together) but please phrase it more carefully.).
            %[suggested sentence: 
            If the templates in Fig.~\ref{fig: 1by2T} truly belonged in two groups, they would be grouped into two groups in this network, even when we try to cluster them into three. 
            However, in the lower panel of Fig.~\ref{fig: 1by3T}, we can see that the middle node contains two templates.
            These two templates, which are separated from the group of starburst templates in the lower part of Fig.~\ref{fig: 1by2T},  are the SB5 and SB6 types.
            In the upper plot of Fig.~\ref{fig: 1by3T}, the colour between two right neurons is black and the colour between two left neurons is white. 
            The black colour indicates that the left neuron is completely different from the other two groups,
            while the white colour shows that the two right neurons are very similar to one another.

@@

            \begin{figure}
                \begin{subfigure}[b]{0.5\textwidth}
                    \centering
                    \includegraphics[width=\textwidth]{../image_paper2/1d/dist_1_by_3.png}
                \end{subfigure}
                \hfill
                \begin{subfigure}[b]{0.5\textwidth}
                     \includegraphics[width=\textwidth]{../image_paper2/1d/hit_t_1_by_3.png}
                \end{subfigure}
                \caption{The same as Fig.~\ref{fig: 1by2T} but showing the results of training network in $1\times3$~grid. In this network again, 5 of the \citet{Kinney96} templates are categorized as quiescent and 7 as starbursts. However, this time 2  templates (SB5 and SB6) are separated from the starburst groups.}
                 \label{fig: 1by3T}
            \end{figure}

            We increased the size of the maps gradually until the galaxies are divided into twelve groups (Figs.~\ref{fig: 1by4T} to ~\ref{fig: 1by20T} and ~\ref{fig: 1by22T}).
            Since there are more nodes in higher grid SOMs, the  algorithm 
            pays more attention to small differences between groups.
            If the templates from \citetalias{Kinney96} had completely distinct SEDs, then a $1\times12$-sized SOM would be expected to show 12 different groups each containing a single template.
            However, in Fig.~\ref{fig: 1by12T}, we see that three of the neurons contain two templates and three of them are empty.
            It is evident that there was no template in the \citetalias{Kinney96} sample that can fill those empty neurons.
            In Fig.~\ref{fig: 1by12T}, from left to right, templates with types B and E, SB3 and SB4, and SB1 and SB2 are the ones grouped together. 
            The SB3 and SB4 grouping breaks when we increase the size of the network to $1\times15$~(Fig.~\ref{fig: 1by15T}).
            The SB1 and SB2 templates, however, remain in the same neuron until the size of the map is increased to $1\times20$~(Fig.~\ref{fig: 1by20T}).
            The separation between templates B and E only happens when the size of the SOM exceeds $1\times22$~(Fig.~\ref{fig: 1by22T}).
        \begin{figure*}
            \begin{subfigure}[b]{\textwidth}
                \centering
                \includegraphics[width=\textwidth]{../image_paper2/1d/dist_1_by_22.png}
            \end{subfigure}
            \hfill
            \begin{subfigure}[b]{\textwidth}
                \includegraphics[width=\textwidth]{../image_paper2/1d/hit_t_1_by_22.png}
            \end{subfigure}
            \caption{The same as Fig.~\ref{fig: 1by2T} but this time the figure shows results of training network in $1\times22$~grid.}
            \label{fig: 1by22T}
        \end{figure*}

@@

        \subsubsection{Classifying the galaxy sample}
         \label{sec: 1Dv}
            After training the networks as discussed in Sec.~\ref{sec: 1Dt}, we used them to classify the fitted SEDs of the sample of 142 galaxies from \citetalias{Hossein12}.
            The upper panel in Fig.~\ref{fig: 1by2V} shows the result of this classification using the $1\times2$~network from Fig.~\ref{fig: 1by2T}.
            Eighty of the galaxies have SEDs similar to those of the early type galaxies and the SEDs of the rest are similar to the starburst galaxies.
            %The lower panels show the average SED of the galaxies in each node. 
            The lower left panel shows the average SED of the 80 galaxies that were classified as early type 
            These galaxies are similar to the ones in the left node in the upper panel of Fig.~\ref{fig: 1by2T}:
            the SEDs clearly show the 4000\AA~break, one of the signatures of early type galaxies.
            The H$\alpha$ emission in the spectrum could be from galaxies with similar SEDs to Sa galaxies, but with stronger emission lines.
            The average SED in the lower right panel of Fig.~\ref{fig: 1by2V} shows strong emission lines and bright ultraviolet continuum, indications of a high star formation rate.
            \begin{figure}
                \begin{subfigure}[b]{0.5\textwidth}
                    \centering
                    \includegraphics[width=\textwidth]{../image_paper2/1d/hit_v_1_by_2.png}
                    %\caption{$1\times2$ weight map}
                     %\label{fig: 1by3T}
                \end{subfigure}
                \hfill
                \begin{subfigure}[b]{0.5\textwidth}
                     \includegraphics[width=\textwidth]{../image_paper2/1d/SED_total1by2.png}
                     %\caption{$1\times2$ hits map}
                     %\label{fig: 1by3Thits}
                \end{subfigure}
                \caption{Classification of fitted galaxy SEDs from \citet{Hossein12} using the $1\times2$~networks trained from the \citet{Kinney96} templates (Fig.~\ref{fig: 1by2T}). Upper panel: a hit map with the number in each node representing the number of galaxies belonging to that group. In this case 80 means that 80 of the 142 SEDs are classified into the early type group while the 62 of the 142 SEDs are categorized as starburst galaxies. Lower panel: Average SED of the galaxies in each group, early-type on the left and starburst on the right.}
                \label{fig: 1by2V}
            \end{figure}

@@

            \begin{figure}
                \begin{subfigure}[b]{0.5\textwidth}
                    \centering
                    \includegraphics[width=\textwidth]{../image_paper2/1d/hit_v_1_by_3.png}
                    %\caption{$1\times2$ weight map}
                     %\label{fig: 1by3T}
                \end{subfigure}
                \hfill
                \begin{subfigure}[b]{0.5\textwidth}
                     \includegraphics[width=\textwidth]{../image_paper2/1d/SED_total1by3.png}
                     %\caption{$1\times2$ hits map}
                     %\label{fig: 1by3Thits}
                \end{subfigure}
                \caption{Same as Fig.~\ref{fig: 1by2V}, but in this figure, we used a network with size of $1\times3$ (Fig.~\ref{fig: 1by3T}) to classify the sample galaxies.}
                \label{fig: 1by3V}
            \end{figure}

>>>>

\chapter[SOM: classifying high $Z$ galaxies]{Clustering galaxy spectra at $0.5<z<1$: an unsupervised approach}
\label{ch: paper2}
\defcitealias{Hossein12}{T12}
\defcitealias{Kinney96}{K96}
\defcitealias{Noll09}{N09}

@@

 \subsection{Kinney spectral model}
     \begin{figure}
        \centering
        \includegraphics[width=\textwidth]{../image_paper2/k96.jpg}
        \caption[\citet{Kinney96} spectra template for 12 types of galaxies]{\citet{Kinney96} spectra template for 12 types of galaxies from \citet{Hossein12} paper (Fig. 1 in \citet{Hossein12}). The type of each template is shown in each frame. Plots B, E, S0, Sa, Sb and SC show spectra that belong to the early types galaxies. Starburst galaxies spectra are indicated with SB 1 to 6. Higher numbers represent more intrinsic colour extinction.}
        \label{fig: k96}
    \end{figure}

@@

 \subsection{SED and properties of the sample galaxies} 
    \citetalias{Hossein12} selected 142 galaxies from the spectroscopic campaign of the ESO GOODS-South field~\citep{Vanzella05, Vanzella06, Vanzella08}.
    The 142 galaxies were selected based on the availability of photometry from HST/ACS, VLA/ISAAC, and {\it Spitzer}/MIPS and IRAC (10--13 filters with $\sim 0.4<\lambda<24~\mu$m in the observed frame).
   Data from these instruments was necessary in order to have a complete picture of stellar population and star formation rate. 
    For each galaxy, a robust spectroscopic redshift and photometric measurements from the GOODS-MUSIC catalogue \citep{Santini09} was available.
   \citetalias{Hossein12} matched the point spread function (PSF) of the photometric data with \citetalias{Kinney96} data and used the photometry as inputs to the Code Investigating GALaxy Emission ({\em CIGALE});~\citep[][hereafter N09]{Noll09} to generate the best-fit SED for each galaxy.
   % as well as some of the physical properties of the galaxies.
% PB 20160603: commented out following 3 lines as I don't think this fit well here. Could maybe go somewhere else.    
%    {\em CIGALE} is a valuable tool to investigate the properties of galaxies using ultraviolet to infrared wavebands.
%    It uses stellar populations, synthetic attenuation and dust emission models, spectral line templates, and active galactic nuclei optical spectral templates to model the SEDs of galaxies.
%    This code was successfully tested on data from 39 galaxies selected from the Spitzer Infrared Nearby Galaxy Survey (SINGS;~\citep{Kennicutt03}) by \citetalias{Noll09}.
    \citetalias{Hossein12} produced the best SED match, with wavelength interval of 910~\AA~to $\sim 80$~cm, for each galaxy\footnote{This wavelength interval is the default output of the {\em CIGALE} code}.
    Assuming decreasing SFR and visual attenuation ($\tau$) model, Salpeter initial mass function~\citep{Salpeter55}, and old stellar population with age of $\sim 10$~Gyr, they derived physical properties of the galaxies such as age and stellar mass.
    Some of these properties are shown in Table~\ref{tab: props}.
    In Section~\ref{sec: 1D}, we study these properties for each category.
    More details on creating SEDs and extracting information about galaxy properties using {\em CIGALE} can be found in \citetalias{Noll09} and \citetalias{Hossein12}.

\begin{table}
\caption[Description of the properties of \citet{Hossein12} galaxies]{Description of the properties of \citet{Hossein12} galaxies; the output result of {\em CIGALE}}     
\label{tab: props}
\centering
\begin{tabular}{l l l}
\hline\hline
\noalign{\smallskip}
Par. & Unit & Description\\
\noalign{\smallskip}
\hline
\noalign{\smallskip}
$t_{\,\mathrm{oSP}}$ & Gyr & age of old SP model \\
$t_{\,\mathrm{ySP}}$ & Gyr & age of young SP model \\
$f_\mathrm{burst}$ & --- & mass fraction of \\
& & young single population (SP) model \\
\noalign{\smallskip}
$t_{\,\mathrm{D4000}}$ & Gyr & D4000-related age \\
\noalign{\smallskip}
$M_\mathrm{star}$ & M$_\odot$ & total stellar mass  \\
SFR & M$_\odot$/yr & instantaneous SFR  \\
$A_\mathrm{FUV}$ & mag & attenuation at 1500\,\AA{} \\
\noalign{\smallskip}
\hline
\end{tabular}
\end{table}

@@

\subsection{Creating self-organizing maps}
\label{sec: create_som}
     In order to create SOMs, we use {\sc matlab} neural network toolbox~\citep[NNT,][]{matlabtolbox}.

     SOM in {\sc nnt} can be created by {\sc newsom} or {\sc selforgmap} library, both of which work in two phases, an ``ordering phase" and a ``tuning phase". 
     The first phase is called the ``ordering phase" and
     starts with maximum neighbourhood distance and an initial high learning factor (usually 0.9) is provided by the user. 
     The ordering phase continues for a requested number of iterations.
     During the iterations, the learning factor decreases to the tuning phase learning factor and the neighbourhood distance reaches that of the tuning phase as well.
     Both the learning factor and the neighbourhood of the tuning phase are set by the user. 
     The amount by which these two factors change in each iteration depends on the number of iterations.

     In the second, or ``tuning'' phase,
     the neighbourhood distance is kept at the user-defined minimum.
     The learning factor, however, decreases gradually.
     The gradual change in the leaning factor helps to fine-tune the topology results, leading to a more stable SOM. 
     To allow the fine tuning, the number of iterations in this phase must be much larger than the that of the ordering phase.
     We chose the number of epochs in the tuning phase to be 3 times the number of epochs in the ordering phase.

     To present our results, we use {\sc nnt}'s built-in plotting tool.
     Specifically, we use two of the plots in this tool: a hits map, which shows the number of times each neuron has became the winner (hits), and a distance map, which shows the distance between those neurons.
     In the maps, the purple hexagonal shapes represent the neurons. 
     The distances in a distance map are shown by the grey cycle colours:
     the darker the colour, the larger the distance between neurons.
     In the hit maps, neurons with zero hits are left empty.

@@

         \begin{figure}
            \begin{subfigure}[b]{0.5\textwidth}
                \centering
                \includegraphics[width=\textwidth]{../image_paper2/sample/sample2_dist.png}
            \end{subfigure}
            \hfill
            \begin{subfigure}[b]{0.5\textwidth}
                \includegraphics[width=\textwidth]{../image_paper2/sample/sample2_hits.png}
            \end{subfigure}
            \caption[Self-organizing map of the mock sample]{SOM of the mock sample. In both plots, the axes show the position of the neurons. Hexagonal shapes represent the neurons. The top plot is a distance map. The grey cycle colours show the differences between the weights of each neuron with white being the minimum difference and black being the maximum one. The lower plot is a ``hits plot". It represents the number of samples in each neuron, where empty means zero hits. In this sample, the 27 galaxies are clustered in 4 groups, containing 8, 7, 6 and 6 galaxies respectively.}
            \label{fig: sample}
        \end{figure}

 To illustrate how self-organizing maps work, we create a mock sample of 27 galaxies.
 The mock sample contains two attributes for each galaxy: the type (type 1 or type 2), and the star formation rate (high or low). 
 We generated a SOM of size $10 \times 10$ using the same initial values mentioned in Section~\ref{sec: create_som}.

@@

Using this method, as expected, we are able to divide the mock sample galaxies into 4 distinct groups: Type 1 with high SFR, type 1 with low SFR, type 2 with high SFR and type 2 with low SFR. 
The upper panel in the Fig.~\ref{fig: sample} clearly shows this division.
In that plot, the upper half belongs to high star forming galaxies, while the lower half belongs to low star forming galaxies.
The left half of the plot is where type 1 galaxies belong, and type 2 galaxies reside in the right side.
Different shades of gray show the border between regions.
The lower panel in Fig.~\ref{fig: sample} shows that only 4 neurons (out of 100) are occupied. 
8 galaxies are type 1 galaxies with high SFR, 7 are type 2 galaxies with high SFR, 6 are type 1 galaxies with low SFR and the other 6 are type 2 galaxies with low SFR. 
We can conclude that although each galaxy had  the chance to occupy any of the neurons in the network, because of the similarity in the values of their attributes, they remained only in four groups.
This network is considered a training network and can be used to cluster any new dataset with similar entries.

@@

    In this section we show the results of the neural networks trained using the \citetalias{Kinney96} template spectra.
    Training with \citetalias{Kinney96} templates results in networks that have regions corresponding to galaxies of known morphological type. 
    The trained networks can then be used to categorize other galaxies.

@@

    We started our analysis by creating 1D SOMs. 
    First, we created SOMs with only two neurons ($1\times2$ map), and then increased the number of neurons one at a time in the 1D case (Section~\ref{sec: 1D}).
    We generate 2D networks (Section~\ref{sec: 2D}),  again starting with the smallest possible number of neurons (4 neurons in a $2\times2$ map), increasing to 144 in a $12\times12$ map.    
    For each generated network, we compare the results with the \citetalias{Kinney96} categorization.
    We also use these networks to classify the \citetalias{Hossein12} galaxy sample, and compare this classification with that from the supervised networks in \citetalias{Hossein12}.

@@

            Figs.~\ref{fig: 1by2T} --\ref{fig: 1by22T} show the results of the training networks.
            As in to Fig.~\ref{fig: sample}, the upper part of the figures shows the neurons and their relative distances between weight of neurons.
            As mentioned in Section~\ref{sec: method}, an increase in the darkness of colours between neurons represents an increase in relative distance between the neurons.
            The lower panels of the figures show the number of \citetalias{Kinney96} templates that are placed in each neuron. 
            \begin{figure}
                \begin{subfigure}[b]{\textwidth}
                    \centering
                    \includegraphics[width=\textwidth]{../image_paper2/1d/dist_1_by_2.png}
                \end{subfigure}
                \hfill
                \begin{subfigure}[b]{\textwidth}
                     \includegraphics[width=\textwidth]{../image_paper2/1d/hit_t_1_by_2.png}
                \end{subfigure}
                \caption[Results of training network in $1\times2$~grid]{Results of training network in $1\times2$~grid. As in Fig.~\ref{fig: sample}, the upper panel is a distance map and the lower panel is a hit map. In this network, 5 of the templates from \citet{Kinney96} are categorized as quiescent galaxies and the rest are starbursts. Because of their strong emission lines, Sc galaxies are moved towards the starburst ones.}
                 \label{fig: 1by2T}
            \end{figure}

@@

            Fig.~\ref{fig: 1by3T} shows the results of the training in a 1$\times$3 network.
            In these plots we force the galaxies to be categorize in a maximum 3 groups. 
            If the templates in Fig.~\ref{fig: 1by2T} truly belonged in two groups, they would be grouped into two groups in this network, even when we try to cluster them into three. 
            However, in the lower panel of Fig.~\ref{fig: 1by3T}, we can see that the middle node contains two templates.
            These two templates, which are separated from the group of starburst templates in the lower part of Fig.~\ref{fig: 1by2T},  are the SB5 and SB6 types.
            In the upper plot of Fig.~\ref{fig: 1by3T}, the colour between two right neurons is black and the colour between two left neurons is white. 
            The black colour indicates that the left neuron is completely different from the other two groups,
            while the white colour shows that the two right neurons are very similar to one another.

@@

            \begin{figure}
                \begin{subfigure}[b]{\textwidth}
                    \centering
                    \includegraphics[width=\textwidth]{../image_paper2/1d/dist_1_by_3.png}
                \end{subfigure}
                \hfill
                \begin{subfigure}[b]{\textwidth}
                     \includegraphics[width=\textwidth]{../image_paper2/1d/hit_t_1_by_3.png}
                \end{subfigure}
                \caption[Results of training network in $1\times3$~grid]{The same as Fig.~\ref{fig: 1by2T} but showing the results of training network in $1\times3$~grid. In this network again, 5 of the \citet{Kinney96} templates are categorized as quiescent and 7 as starbursts. However, this time 2  templates (SB5 and SB6) are separated from the starburst groups.}
                 \label{fig: 1by3T}
            \end{figure}

            We increased the size of the maps gradually until the galaxies are divided into twelve groups (Figs.~\ref{fig: 1by4T} to ~\ref{fig: 1by20T} and ~\ref{fig: 1by22T}).
            Since there are more nodes in higher grid SOMs, the  algorithm 
            pays more attention to small differences between groups.
            If the templates from \citetalias{Kinney96} had completely distinct SEDs, then a $1\times12$-sized SOM would be expected to show 12 different groups each containing a single template.
            However, in Fig.~\ref{fig: 1by12T}, we see that three of the neurons contain two templates and three of them are empty.
            It is evident that there was no template in the \citetalias{Kinney96} sample that can fill those empty neurons.
            In Fig.~\ref{fig: 1by12T}, from left to right, templates with types B and E, SB3 and SB4, and SB1 and SB2 are the ones grouped together. 
            The SB3 and SB4 grouping breaks when we increase the size of the network to $1\times15$~(Fig.~\ref{fig: 1by15T}).
            The SB1 and SB2 templates, however, remain in the same neuron until the size of the map is increased to $1\times20$~(Fig.~\ref{fig: 1by20T}).
            The separation between templates B and E only happens when the size of the SOM exceeds $1\times22$~(Fig.~\ref{fig: 1by22T}).
        \begin{figure}
            \begin{subfigure}[b]{\textwidth}
                \centering
                \includegraphics[width=\textwidth]{../image_paper2/1d/dist_1_by_22.png}
            \end{subfigure}
            \hfill
            \begin{subfigure}[b]{\textwidth}
                \includegraphics[width=\textwidth]{../image_paper2/1d/hit_t_1_by_22.png}
            \end{subfigure}
            \caption[Results of training network in $1\times22$~grid]{The same as Fig.~\ref{fig: 1by2T} but this time the figure shows results of training network in $1\times22$~grid.}
            \label{fig: 1by22T}
        \end{figure}

@@

        \subsubsection{Classifying the galaxy sample}
         \label{sec: 1Dv}
            After training the networks as discussed in Section~\ref{sec: 1Dt}, we used them to classify the fitted SEDs of the sample of 142 galaxies from \citetalias{Hossein12}.
            The upper panel in Fig.~\ref{fig: 1by2V} shows the result of this classification using the $1\times2$~network from Fig.~\ref{fig: 1by2T}.
            Eighty of the galaxies have SEDs similar to those of the early type galaxies and the SEDs of the rest are similar to the starburst galaxies.
            The lower left panel shows the average SED of the 80 galaxies that were classified as early type 
            These galaxies are similar to the ones in the left node in the upper panel of Fig.~\ref{fig: 1by2T}:
            the SEDs clearly show the 4000\AA~break, one of the signatures of early type galaxies.
            The H$\alpha$ emission in the spectrum could be from galaxies with similar SEDs to Sa galaxies, but with stronger emission lines.
            The average SED in the lower right panel of Fig.~\ref{fig: 1by2V} shows strong emission lines and bright ultraviolet continuum, indications of a high star formation rate.
            \begin{figure}
                \begin{subfigure}[b]{\textwidth}
                    \centering
                    \includegraphics[width=\textwidth]{../image_paper2/1d/hit_v_1_by_2.png}
                    %\caption{$1\times2$ weight map}
                     %\label{fig: 1by3T}
                \end{subfigure}
                \hfill
                \begin{subfigure}[b]{\textwidth}
                     \includegraphics[width=\textwidth]{../image_paper2/1d/SED_total1by2.png}
                     %\caption{$1\times2$ hits map}
                     %\label{fig: 1by3Thits}
                \end{subfigure}
                \caption[Classification of fitted galaxy SEDs from \citet{Hossein12} using the $1\times2$~networks]{Classification of fitted galaxy SEDs from \citet{Hossein12} using the $1\times2$~network trained from the \citet{Kinney96} templates (Fig.~\ref{fig: 1by2T}). Upper panel: a hit map with the number in each node representing the number of galaxies belonging to that group. In this case 80 means that 80 of the 142 SEDs are classified into the early type group while the 62 of the 142 SEDs are categorized as starburst galaxies. Lower panel: Average SED of the galaxies in each group, early-type on the left and starburst on the right.}
                \label{fig: 1by2V}
            \end{figure}

@@

            \begin{figure}
                \begin{subfigure}[b]{\textwidth}
                    \centering
                    \includegraphics[width=\textwidth]{../image_paper2/1d/hit_v_1_by_3.png}
                \end{subfigure}
                \hfill
                \begin{subfigure}[b]{\textwidth}
                     \includegraphics[width=\textwidth]{../image_paper2/1d/SED_total1by3.png}
                \end{subfigure}
                \caption[Classification of fitted galaxy SEDs from \citet{Hossein12} using the $1\times3$~networks]{Same as Fig.~\ref{fig: 1by2V}, but in this figure, we used a network with size of $1\times3$ (Fig.~\ref{fig: 1by3T}) to classify the sample galaxies.}
                \label{fig: 1by3V}
            \end{figure}

            In Fig.~\ref{fig: 1by22V}, we use the $1\times22$~network to classify the sample galaxies.
            As mentioned in Section~\ref{sec: 1Dt}, in this network size we observed the first separation of the \citetalias{Kinney96} galaxies into 12 different neurons.
           As in Figs.~\ref{fig: 1by2V} and ~\ref{fig: 1by3V}, the upper panel of Fig.~\ref{fig: 1by22V} shows the number of galaxies (out of the 142) belonging to each neuron in the $1\times22$ SOM.

            \begin{figure}
                \begin{subfigure}[b]{\textwidth}
                    \includegraphics[width=\textwidth]{../image_paper2/1d/hit_v_1_by_22_n.png}
                    \centering
                \end{subfigure}
                \hfill
                \begin{subfigure}[b]{\textwidth}
                \centering
                     \includegraphics[width=\textwidth]{../image_paper2/1d/SED_total1by22.png}
                \end{subfigure}
                \caption[Classification of fitted galaxy SEDs from \citet{Hossein12} using the $1\times22$~networks]{Same as Fig.~\ref{fig: 1by2V}, but in this figure, we used a network with size of $1\times22$ (Fig.~\ref{fig: 1by22T}) to classify the sample galaxies. In the average SED of each neuron, we can clearly see the changes from early type galaxies to the starburst ones. The 4000\AA~ break weakens from left to right and the emission line strengths increase.}
                \label{fig: 1by22V}
            \end{figure}

            The lower panels present the average SED of the galaxies in each neuron.
            Since galaxies had more space to separate, some of the neurons are left empty.
            Thus, instead of having 22 average SED plots in the lower part of Fig.~\ref{fig: 1by22V}, there are only 12.
